Rclean: A Tool for Writing Cleaner, More Transparent Code

### 2020-02-04 joss/paper.Rmd
Spellcheck
renders pdf and md

In the process of cleaning a script, it is likely that a user will not
know the precise objects they want and might require some help
analyzing it. There are several functions to help with this
process. The `get_vars` function will return a list of possible
results for a given script at a supplied file path. This is obviously
an important step, and justifiably, the default behavior of the
`clean` function is to run `get_vars` if no results are supplied. To
help with limiting and checking the selection of results, the
`code_graph` function creates a network graph of the relationships
among the various results and lines of code in the script, which is
the function used to create the figure demonstrating data provenance
\ref{prov-graph}. Last, the `get_libs` function can be used to detect
the packages that a given script depends on, which it will return as
coded library calls that can be inserted into a cleaned script.

@@

There are multiple approaches to collecting data provenance, but
`Rclean` uses "prospective" provenance, which analyzes code and uses
language specific information to predict the relationship among
processes and data objects. `Rclean` relies on a library called
`CodeDepends` to gather the prospective provenance for each
script. For more information on the mechanics of the `CodeDepends`
package, see [@Lang2019]. To get an idea of what data provenance is,
take a look at the `code_graph` function. The plot that it generates
is a graphical representation of the prospective provenance generated
for `Rclean` \ref{prov-graph}.

@@

Although the workflow described is impactful, we see promise in
extending it to interface with other clean code and reproducibility
tools. One example is the `reprex` package, which provides a simple
API for sharing reproducible examples. Another possibility is to help
transtion scripts to function, package and workflow creation and
refactoring via toolboxes like `drake`. `Rclean` could provide a
reliable way to extract parts of a larger script that would be piped
to a simplified reproducible example, in the case of `reprex`, or,
since it can isolate the code from inputs to one or more outputs, be
used to extract all of the components needed to write one or more
functions that would be a part of a package or workflow, as is the
goal of `drake`.

>>>>

In the process of cleaning a script, it is likely that a user will not
know the precise objects they want and might require some help
analyzing it. There are several functions to help with this
process. The `get_vars` function will return a list of possible
results for a given script at a supplied file path. This is obviously
an important step, and justifiably, the default behavior of the
`clean` function is to run `get_vars` if no results are supplied. To
help with limiting and checking the selection of results, the
`code_graph` function creates a network graph of the relationships
among the various results and lines of code in the script, which is
the function used to create the figure demonstrating data provenance
\@ref(fig:prov-graph). Last, the `get_libs` function can be used to detect
the packages that a given script depends on, which it will return as
coded library calls that can be inserted into a cleaned script.

@@

There are multiple approaches to collecting data provenance, but
`Rclean` uses "prospective" provenance, which analyzes code and uses
language specific information to predict the relationship among
processes and data objects. `Rclean` relies on a library called
`CodeDepends` to gather the prospective provenance for each
script. For more information on the mechanics of the `CodeDepends`
package, see [@Lang2019]. To get an idea of what data provenance is,
take a look at the `code_graph` function. The plot that it generates
is a graphical representation of the prospective provenance generated
for `Rclean` \@ref(fig:prov-graph).

@@

Although the workflow described is impactful, we see promise in
extending it to interface with other clean code and reproducibility
tools. One example is the `reprex` package, which provides a simple
API for sharing reproducible examples. Another possibility is to help
transition scripts to function, package and workflow creation and
refactoring via toolboxes like `drake`. `Rclean` could provide a
reliable way to extract parts of a larger script that would be piped
to a simplified reproducible example, in the case of `reprex`, or,
since it can isolate the code from inputs to one or more outputs, be
used to extract all of the components needed to write one or more
functions that would be a part of a package or workflow, as is the
goal of `drake`.
